
# Multivariate ANOVA

## Multivariate analysis of variance


* Standard ANOVA has just one response variable.

* What if you have more than one response?

* Try an ANOVA on each response separately.

* But might miss some kinds of interesting dependence between the responses that distinguish the groups.



## Packages
```{r }
library(car)
library(tidyverse)
```

   


## Small example


* Measure yield and seed weight of plants grown under 2 conditions: low and high amounts of fertilizer.

* Data (fertilizer, yield, seed weight):
```{r }
my_url <- "http://www.utsc.utoronto.ca/~butler/d29/manova1.txt"
hilo <- read_delim(my_url, " ")
```
def 

* 2 responses, yield and seed weight.



## The data
```{r }
hilo
```

   


## Boxplot for yield for each fertilizer group
```{r ferto,size="small",fig.height=3}
ggplot(hilo, aes(x = fertilizer, y = yield)) + geom_boxplot()
```
def 
Yields overlap for fertilizer groups.


## Boxplot for weight for each fertilizer group
```{r casteldisangro,size="small",fig.height=3}
ggplot(hilo, aes(x = fertilizer, y = weight)) + geom_boxplot()
```
def 

Weights overlap for fertilizer groups.


## ANOVAs for yield and weight

{\small
```{r }
hilo.y <- aov(yield ~ fertilizer, data = hilo)
summary(hilo.y)
hilo.w <- aov(weight ~ fertilizer, data = hilo)
summary(hilo.w)
```
def 
}

Neither response depends significantly on fertilizer. But\ldots


## Plotting both responses at once

Have two response variables (not more), so can plot the
response variables against *each other*, labelling points by
which fertilizer group they're from.

```{r }
g <- ggplot(hilo, aes(
  x = yield, y = weight,
  colour = fertilizer
)) + geom_point()
```

 

Want line through points $(31,14)$ and $(38,10)$ (why? Later):

```{r size="footnotesize"}
d <- tribble(
  ~line_x, ~line_y,
  31, 14,
  38, 10
)
g <- g + geom_line(data = d, aes(
  x = line_x, y = line_y,
  colour = NULL
))
```

 

\begin{small}
Fitting regression line through points in `d`. 
Adding to previous `ggplot`, so `geom_smooth` 
inherits `colour` from first one. This data frame has
no `fertilizer` (previous `colour`), so have to unset.
\end{small}


## The plot
```{r charlecombe,fig.height=3.5}
g
```
def 


## MANOVA
```{r fig.height=2.5}
g
```

   

\begin{small}


* High-fertilizer plants have both yield and weight high.

* True even though no sig difference in yield or weight individually.

* Drew line separating highs from lows on plot.

\end{small}


## MANOVA finds multivariate differences


* Is difference found by diagonal line significant? MANOVA finds out.
```{r size="footnotesize"}
response <- with(hilo, cbind(yield, weight))
hilo.1 <- manova(response ~ fertilizer, data = hilo)
summary(hilo.1)
```
def 

* Yes! Difference between groups is *diagonally*, not just up/down
(weight) or left-right (yield). The *yield-weight combination* matters.




## Strategy



* Create new response variable by gluing together columns of
responses, using `cbind`.

* Use `manova` with new response, looks like `lm` otherwise.

* With more than 2 responses, cannot draw graph. What then?

* If MANOVA test significant, cannot use Tukey. What then?

* Use {\em discriminant analysis} (of which more later).




## Another way to do MANOVA
Install (once) and load package `car`:
```{r eval=F}
library(car)
```
def 


## Another way\ldots

```{r size="footnotesize"}
hilo.2.lm <- lm(response ~ fertilizer, data = hilo)
hilo.2 <- Manova(hilo.2.lm)
hilo.2
```

   


* Same result as small-m `manova`.

* `Manova` will also do *repeated measures*, coming up
later.




## Another example: peanuts


*  Three different varieties
of peanuts (mysteriously, 5, 6 and 8) planted in two different
locations.

* Three response variables: `y`, `smk` and
`w`.

```{r message=F,size="footnotesize"}
my_url <- "http://www.utsc.utoronto.ca/~butler/d29/peanuts.txt"
(peanuts.orig <- read_delim(my_url, " "))
```
def 


## Setup for analysis

```{r }
peanuts <- peanuts.orig %>%
  mutate(
    location = factor(location),
    variety = factor(variety)
  )
response <- with(peanuts, cbind(y, smk, w))
head(response)
```
def 


## Analysis (using `Manova)`

```{r size="scriptsize"}
peanuts.1 <- lm(response ~ location * variety, data = peanuts)
peanuts.2 <- Manova(peanuts.1)
peanuts.2
```
def 



* Interaction not quite significant, but main effects are.

* Combined response variable `(y,smk,w)` definitely depends
on location and on variety

* Weak dependence of `(y,smk,w)` on the location-variety *combination.*

* Understanding that dependence beyond our scope right now.


```{r echo=F, warning=F}
pkgs <- names(sessionInfo()$otherPkgs)
pkgs <- paste("package:", pkgs, sep = "")
x <- lapply(pkgs, detach, character.only = TRUE, unload = TRUE)
```

   


