\documentclass[hyperref={pdfpagelabels=false}]{beamer}

\usepackage{graphicx}

% compile with latex and then dvips -t letter

%\setlength{\parindent}{0em}
%\setlength{\parskip}{30pt}

%\setlength{\textwidth}{5in}
%\setlength{\textheight}{5in}

%\setcounter{secnumdepth}{0}


\begin{document}

\begin{frame}{Outline part 1}
\begin{itemize}
    \item  Lecture: Wednesday 14:00-16:00, IC 328
    \item  Instructor: Ken Butler
    \item  Office: IC 471
    \item  E-mail: \verb-butler@utsc.utoronto.ca-
    \item  Office hours: find me on Mon or Wed, make an appointment. E-mail will reach me at any time if you need to get in touch.
    \item Website: http://www.utsc.utoronto.ca/~butler/d29/
    \item Blackboard: grades (and maybe for mass e-mails). Otherwise,
      see website.
\end{itemize}
\end{frame}

\begin{frame}{Outline part 2}
\begin{itemize}
    \item {\bf No official text} for this course. One that shares my philosophy:
      \begin{itemize}
      \item ``Using Multivariate Statistics'' by Barbara G.\ Tabachnick and Linda S.\ Fidell, publ.\ Allyn and Bacon, ISBN 0205459382. There is a 5th edition around, but the 4th should be fine.
      \end{itemize}

    \item Prerequisites: for undergrads STAB22 - STAB27
      (or STA220-STA221). For grad students,
      a first course, and some training in
      regression and ANOVA.
    \item  Exclusions: this course not for Statistics majors. For students in other fields who wish to learn some more advanced statistical methods. The exclusions in the Calendar reflect this.
\end{itemize}
\end{frame}

\begin{frame}{Topics (mutable)}

{\small
      \begin{itemize}
          \item  review of statistical inference; two-sample t-test
          \item  review of multiple regression
          \item  logistic regression, including ordinal and nominal response
          \item survival analysis
          \item a brief review of analysis of variance
          \item analysis of covariance; multivariate analysis of variance
          \item repeated measures (by profile analysis)
          \item  discriminant analysis; cluster analysis
          \item multidimensional scaling
          \item  principal components
          \item  factor analysis (exploratory and confirmatory)
          \item time series analysis (an introduction)
          \item nonparametric statistics
          \item frequency tables
          \end{itemize}
}

Let me know if there's something else you'd like to see.


\end{frame}

\begin{frame}{Computing and assessment}
\begin{itemize}
\item Computing: big part of the course, {\bf not}
  optional. Demonstrate that you can use 
  SAS to analyze data, and to
  critically interpret the output. {\em No} prior knowledge of SAS is assumed.
\item Grading: (3 hour) final exam, but no midterm. There will
  be assignments most weeks. 
  Graduate students also required to
  complete a project using one or more of the techniques learned in
  class, on a dataset from their field of study.  For 
\item Undergrads:
  final exam 30\% and assignments (equally weighted)
  70\%; grad students: the final exam also 30\%, the
  assignments 35\% and project 35\%.
  Projects due at the
  last class of the semester.
    \item Plagiarism: don't do it!
\end{itemize}
\end{frame}


\end{document}






