% $Header: /home/vedranm/bitbucket/beamer/solutions/conference-talks/conference-ornate-20min.en.tex,v 90e850259b8b 2007/01/28 20:48:30 tantau $

\documentclass[unknownkeysallowed]{beamer}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
\let\hlipl\hlkwb

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}



\mode<presentation>
{
  \usetheme{Warsaw}
  \setbeamertemplate{headline}{}

  % or ...

  \setbeamercovered{transparent}
  % or whatever (possibly just delete it)
  \setbeamertemplate{footline}[frame number]
  \setbeamertemplate{navigation symbols}{}%remove navigation symbols
}


\usepackage[english]{babel}
% or whatever

\usepackage[utf8]{inputenc}
\DeclareUnicodeCharacter{00D7}{$\times$}
\DeclareUnicodeCharacter{221A}{\checkmark}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\usepackage{graphicx}
%\usepackage{hyperref}
\usepackage{lastpage}
%\usepackage{Sweave}
\usepackage{verbatim}
\usepackage{multicol}
\usepackage[user]{zref}

\title[Course Slides] % (optional, use only with long paper titles)
{STAD29 / STA1007}
\subtitle
{Statistics for the Life and Social Sciences}

\author % (optional, use only with lots of authors)
{Ken Butler}
% - Give the names in the same order as the appear in the paper.
% - Use the \inst{?} command only if the authors have different
%   affiliation.

%\institute[testing] % (optional, but mostly needed)
%{
%  \inst{1}%
%  Department of Computer and Mathematical Sciences\\
%  University of Toronto Scarborough
%}
%% - Use the \inst command only if there are several affiliations.
% - Keep it simple, no one is interested in your street address.

\date[Winter 2018] % (optional, should be abbreviation of conference name)
{Winter semester 2018}
% - Either use conference name or its abbreviation.
% - Not really informative to the audience, more for people (including
%   yourself) who are reading the slides online

%\subject{Theoretical Computer Science}
% This is only inserted into the PDF information catalog. Can be left
% out. 
 


% If you have a file called "university-logo-filename.xxx", where xxx
% is a graphic format that can be processed by latex or pdflatex,
% resp., then you can add a logo as follows:

% \pgfdeclareimage[height=0.5cm]{university-logo}{university-logo-filename}
% \logo{\pgfuseimage{university-logo}}


% Delete this, if you do not want the table of contents to pop up at
% the beginning of each subsection:
\AtBeginSubsection[]
{
  \begin{frame}<beamer>{Outline}
    \tableofcontents[currentsection,currentsubsection]
  \end{frame}
}


% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command: 

%\beamerdefaultoverlayspecification{<+->}

% my chapters

%\input{preamb}

%\includeonly{bPrincomp}
%\includeonly{bOutline,bRIntro,bggplot}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}
%\setkeys{Gin}{width=0.6\textwidth}


\begin{frame}
  \titlepage
\end{frame}

%\include{bOutline}


\section*{Course Outline}
\frame{\sectionpage}

\begin{frame}[fragile]{Course and instructor}
  \begin{itemize}
    \item  Lecture: Wednesday 14:00-16:00 in HW 215. Optional computer
      lab Monday 16:00-17:00 in BV 498.
    \item  Instructor: Ken Butler
    \item  Office: IC 471.
    \item  E-mail: \verb-butler@utsc.utoronto.ca-
    \item Office hours: Monday 11:00-13:00. I am often around
      otherwise. See if I'm in. Or make an appointment. E-mail always
      good.
    \item Course website: 
\url{www.utsc.utoronto.ca/~butler/d29}.
    \item Using Quercus for assignments/grades only; using website for
      everything else.
\end{itemize}

\end{frame}

\begin{frame}[fragile]{Texts}

\begin{itemize}
\item There is no official text for this course. 
\item You may find ``R for Data Science'', 
  \url{http://r4ds.had.co.nz/} helpful for R background.
\item I will refer frequently to my book of Problems and Solutions in Applied Statistics (PASIAS), \url{http://ritsokiguess.site/pasias/}.
\item Both of these resources are and will remain free.
\end{itemize}
  
\end{frame}

\begin{frame}[fragile]{Programs, prerequisites and exclusions}

  \begin{itemize}
  \item Prerequisites:
      \begin{itemize}
      \item For undergrads: STAC32. Not negotiable.
    \item  For grad students,
      a first course in statistics, and some training in
      regression and ANOVA. The less you know, the more you'll have to
      catch up!
      \end{itemize}

    \item This course is a required part of Applied Statistics minor.
    \item Exclusions: \textbf{this course is not for
        Math/Statistics/CS majors/minors}. It is for students in other
      fields who wish to learn some more advanced statistical
      methods. The exclusions in the Calendar reflect this. 
    \item If you
      are in one of those programs, you won't get program credit for
      this course, \textbf{or for any future STA courses you take.}

  \end{itemize}
  
\end{frame}


\begin{frame}[fragile]{Computing}

  \begin{itemize}
  \item Computing: big part of the course, {\bf not}
    optional. Demonstrate that you can use 
    R to analyze data, and can
    critically interpret the output.
  \item For grad students who have not come through STAC32, I am happy
    to offer extra help to get you up to speed.
  \end{itemize}
  
\end{frame}


\begin{frame}{Assessment}
\begin{itemize}
\item Grading: (2 hour) midterm, (3 hour) final exam. Assignments most
  weeks, due Tuesday at 11:59pm. 
  Graduate students (STA 1007) also required to
  complete a project using one or more of the techniques learned in
  class, on a dataset from their field of study.    Projects due on
  the last day of classes.

\item Assessment:

  \begin{tabular}{lcc}
    & STAD29 & STA 1007\\
    Assignments & 20\% & 20\%\\
    Midterm exam & 30\%  & 20\% \\
    Project & - & 20\%\\
    Final exam & 50\% & 40\%
  \end{tabular}

\item Assessments missed \emph{with documentation} will cause a
  re-weighting of other assessments of same type. No make-ups.
\item You \textbf{must pass the final exam} to pass the course. If you
  fail the final exam but would otherwise have passed the course, you
  receive a grade of 45.

\end{itemize}
\end{frame}

\begin{frame}{Plagiarism}

  \begin{itemize}
  \item
    \url{http://www.utoronto.ca/academicintegrity/academicoffenses.html}
    defines academic offences at this university. Read it.
  \item Plagiarism is defined (at the end) as
    \begin{quote}
       The wrongful appropriation and purloining, and publication as one’s own, of the ideas, or the expression of the ideas ... of another.
    \end{quote}
    \item The code and
    explanations  that
    you write and hand in must be \emph{yours and yours
      alone}. 
    \item When you hand in work, it is implied that it is
    \emph{your} work. Handing in work, with your name on it, that was actually done by
    someone else is an \emph{academic offence}.
  \item If I am suspicious
    that anyone's work is plagiarized, I will take action.
    
  \end{itemize}
  
\end{frame}

\begin{frame}[fragile]{Getting help}

  \begin{itemize}
  \item The English Language Development Centre supports all students
    in developing better Academic English and critical thinking skills
    needed in academic communication. Make use of the personalized
    support in academic writing skills development. Details and sign-up information:
    \url{http://www.utsc.utoronto.ca/eld/}.
  \item      Students with diverse learning styles and needs are welcome in this
course. In particular, if you have a disability/health consideration
that may require accommodations, please feel free to approach the AccessAbility Services Office as soon as possible. I will
work with you and AccessAbility Services to ensure you can achieve
your learning goals in this course. Enquiries are confidential. The
UTSC AccessAbility Services staff are available by
appointment to assess specific needs, provide referrals and arrange
appropriate accommodations: (416) 287-7560 or by e-mail: \texttt{ability@utsc.utoronto.ca}.


  \end{itemize}
  
\end{frame}


% \begin{frame}{What we (might) cover, part 1}
%   \tableofcontents[sections={1-10}]  % You might wish to add the option [pausesections]
% \end{frame}
% 
% \begin{frame}{What we (might) cover, part 2}
%   \tableofcontents[sections={11-}]  % You might wish to add the option [pausesections]
% \end{frame}
% 
% % skip these
% % <<intro,child="bRIntro.Rnw">>=
% % @ 
% % 
% % <<ggplot,child="bggplot.Rnw">>=
% % @ 
% % 
% % <<Inference,child="bInference.Rnw">>=
% % @
% 
% % start here
% 
% <<Rscript, child="bRscript.Rnw">>=
% @  
% 
% <<Regression, child="bRegression.Rnw">>=
% @ 
% 
% 
% <<bLogistic, child="bLogistic.Rnw">>=
%  
% 
% <<bSurvival, child="bSurvival.Rnw">>=
% @ 
% 
% <<bAnova, child="bAnova.Rnw">>=
% @ 
% 
% <<ancova, child="bAncova.Rnw">>=
% @ 
% 
% <<bManova, child="bManova.Rnw">>=
% @ 
% 
% <<bProfile, child="bProfile.Rnw">>=
% @ 
% 
% <<bDiscrim, child="bDiscrim.Rnw">>=
% @ 
% 
% <<bCluster, child="bCluster.Rnw">>=
% @ 
% 
% <<bMDS, child="bMDS.Rnw">>=
% @ 
% 
% <<bPrincomp, child="bPrincomp.Rnw">>=
% @ 
% 
% <<bFactor, child="bFactor.Rnw">>=
% @ 
%  
% % % <<bTimeSeries, child="bTimeSeries.Rnw">>=
% % % @ 
% 
% <<bMultiway, child="bMultiway.Rnw">>=
% @
% %  

%
%\include{bRIntro}
%\include{bggplot} 
%\include{bInference}
%\include{bRegression}
%\include{bRscript}
%\include{bLogistic}
%\include{bSurvival}
%\include{bAnova}
%\include{bAncova}
%\include{bManova}
%\include{bProfile}
%\include{bDiscrim}
%\include{bCluster}
%\include{bMDS}
%\include{bPrincomp}
%\include{bFactor} % lavaan for confirmatory?
%%\include{bKriging}
%\include{bTimeSeries}
%\include{bMultiway} % poisson glm model? http://ww2.coastal.edu/kingw/statistics/R-tutorials/loglin.html
%\section{Time series}
%\section{Nonparametric statistics}
  
%\include{bSummary}
  


% % \begin{frame}{The end}
% 
% {\centering
% {\Huge
% \hfill
% DONE!
% \hfill
% }
% }
% \end{frame}


\end{document}


